19Q2

### 2019-09-03 19Q2.Rmd
CRISPR GE/CN correlations

```{r drug_spear_calc, warning=FALSE, eval=FALSE}
# CCLE
ccle_filt <- filter(ccle, Drug %in% g2p_drugs & Gene %in% g2p_genes)
ccle_filt$Gene <- factor(ccle_filt$Gene, ordered = FALSE)
ccle_filt <- unfactorize(ccle_filt)
ccle_ge_res <- ccle_filt %>% group_by(Drug, Gene, Drug_Gene) %>% drop_na(TPM) %>% summarize(GE_r = cor.test(y = Score, x = TPM, method = "spearman", use = "na.or.complete")$estimate, GE_p = cor.test(y = Score, x = TPM, method = "spearman", use = "na.or.complete")$p.value)
ccle_ge_res$InG2P <- ifelse(ccle_ge_res$Drug_Gene %in% unique(as.character(g2p_expr$Drug_Gene[!is.na(g2p_expr$Drug_Gene)])), "Yes", "No")
ccle_ge_res$Dataset <- "CCLE"
ccle_cn_res <- ccle_filt %>% group_by(Drug, Gene, Drug_Gene) %>% drop_na(CN) %>% summarize(CN_r = cor.test(y = Score, x = CN, method = "spearman", use = "na.or.complete")$estimate, CN_p = cor.test(y = Score, x = CN, method = "spearman", use = "na.or.complete")$p.value)
ccle_cn_res$InG2P <- ifelse(ccle_cn_res$Drug_Gene %in% unique(as.character(g2p_amp$Drug_Gene[!is.na(g2p_amp$Drug_Gene)])), "Yes", "No")
ccle_cn_res$Dataset <- "CCLE"

# CTRP
ctrp_filt <- filter(ctrp, Drug %in% g2p_drugs & Gene %in% g2p_genes)
ctrp_filt$Gene <- factor(ctrp_filt$Gene, ordered = FALSE)
ctrp_filt <- unfactorize(ctrp_filt)
ctrp_ge_res <- ctrp_filt %>% group_by(Drug, Gene, Drug_Gene) %>% drop_na(TPM) %>% summarize(GE_r = cor.test(y = Score, x = TPM, method = "spearman", use = "na.or.complete")$estimate, GE_p = cor.test(y = Score, x = TPM, method = "spearman", use = "na.or.complete")$p.value)
ctrp_ge_res$InG2P <- ifelse(ctrp_ge_res$Drug_Gene %in% unique(as.character(g2p_expr$Drug_Gene[!is.na(g2p_expr$Drug_Gene)])), "Yes", "No")
ctrp_ge_res$Dataset <- "CTRP"
ctrp_cn_res <- ctrp_filt %>% group_by(Drug, Gene, Drug_Gene) %>% drop_na(CN) %>% summarize(CN_r = cor.test(y = Score, x = CN, method = "spearman", use = "na.or.complete")$estimate, CN_p = cor.test(y = Score, x = CN, method = "spearman", use = "na.or.complete")$p.value)
ctrp_cn_res$InG2P <- ifelse(ctrp_cn_res$Drug_Gene %in% unique(as.character(g2p_amp$Drug_Gene[!is.na(g2p_amp$Drug_Gene)])), "Yes", "No")
ctrp_cn_res$Dataset <- "CTRP"

# GDSC
gdsc_filt <- filter(gdsc, Drug %in% g2p_drugs & Gene %in% g2p_genes)
gdsc_filt$Gene <- factor(gdsc_filt$Gene, ordered = FALSE)
gdsc_filt <- unfactorize(gdsc_filt)
gdsc_ge_res <- gdsc_filt %>% group_by(Drug, Gene, Drug_Gene) %>% drop_na(TPM) %>% summarize(GE_r = cor.test(y = Score, x = TPM, method = "spearman", use = "na.or.complete")$estimate, GE_p = cor.test(y = Score, x = TPM, method = "spearman", use = "na.or.complete")$p.value)
gdsc_ge_res$InG2P <- ifelse(gdsc_ge_res$Drug_Gene %in% unique(as.character(g2p_expr$Drug_Gene[!is.na(g2p_expr$Drug_Gene)])), "Yes", "No")
gdsc_ge_res$Dataset <- "GDSC"
gdsc_cn_res <- gdsc_filt %>% group_by(Drug, Gene, Drug_Gene) %>% drop_na(CN) %>% summarize(CN_r = cor.test(y = Score, x = CN, method = "spearman", use = "na.or.complete")$estimate, CN_p = cor.test(y = Score, x = CN, method = "spearman", use = "na.or.complete")$p.value)
gdsc_cn_res$InG2P <- ifelse(gdsc_cn_res$Drug_Gene %in% unique(as.character(g2p_amp$Drug_Gene[!is.na(g2p_amp$Drug_Gene)])), "Yes", "No")
gdsc_cn_res$Dataset <- "GDSC"

@@

```{r load_drug_spear}
g2p_amp_4merge <- select(g2p_amp, Drug_Gene, Direction_Curated)
cn_cor <- readRDS("./data_munging/rds/drug_cn_cor.rds")
cn_cor <- merge(cn_cor, g2p_amp_4merge, by = "Drug_Gene", all.x = TRUE)

@@

Spearman correlations for gene expression and copy number:
```{r crispr_spear_calc, warning=FALSE}
crispr_spear_ge <- transform(crispr %>% group_by(Gene) %>% drop_na(TPM) %>% summarize(p = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$p.value, p_log10 = -log10(p), corr = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$estimate, Metric = "Gene Expression"), Gene = reorder(Gene, corr))
crispr_spear_cn <- transform(crispr %>% group_by(Gene) %>% drop_na(CN) %>% summarize(p = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$p.value, p_log10 = -log10(p), corr = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$estimate, Metric = "Copy Number"), Gene = reorder(Gene, corr))
c("white", "deepskyblue4")

crispr_spear_cn_plot <- ggplot(data = filter(crispr_spear_cn, Gene %in% g2p_dg_genes)) +
  geom_point(aes(x = Gene, y = corr, fill = p_log10), pch = 21, size = 5) +
  theme(legend.position = "bottom") +
  scale_fill_gradientn(colours = c("white", "green4"),
                       breaks = c(0, 4, 8, 12, 16, 20),
                       labels = c(0, 4, 8, 12, 16, 20),
                       limits = c(0, 21)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10)) +
  labs(title = "Copy Number",
       fill = expression(-log[10](p)),
       x = "Gene",
       y = expression("Spearman Correlation Coefficient"~(r[s])))

@@

```{r corr_heatmap_function}
MakeCorrHeatmap <- function(drug_cor_res, row_names, col_title, color_range, legend_title, metric, radius) {
  if(toupper(metric) == "CN" | toupper(metric) == "COPY NUMBER") {
    signif_grid <- dcast(drug_cor_res, Gene ~ Drug, value.var = "CN_p", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Gene)
    signif_grid <- signif_grid[g2p_gene_order, intersect(g2p_drug_order, colnames(signif_grid))]
    signif_grid <- signif_grid[complete.cases(signif_grid), ]
    signif_grid[signif_grid >= 0.05] <- ""
    signif_grid[signif_grid != ""] <- "*"

    corr_grid <- dcast(drug_cor_res, Gene ~ Drug, value.var = "CN_r", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Gene)
    corr_grid <- corr_grid[g2p_gene_order, intersect(g2p_drug_order, colnames(corr_grid))]
    corr_grid <- corr_grid[complete.cases(corr_grid), ]
  }

  if(toupper(metric) == "GE" | toupper(metric) == "GENE EXPRESSION") {
    signif_grid <- dcast(drug_cor_res, Gene ~ Drug, value.var = "GE_p", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Gene)
    signif_grid <- signif_grid[g2p_gene_order, intersect(g2p_drug_order, colnames(signif_grid))]
    signif_grid <- signif_grid[complete.cases(signif_grid), ]
    signif_grid[signif_grid >= 0.05] <- ""
    signif_grid[signif_grid != ""] <- "*"

    corr_grid <- dcast(drug_cor_res, Gene ~ Drug, value.var = "GE_r", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Gene)
    corr_grid <- corr_grid[g2p_gene_order, intersect(g2p_drug_order, colnames(corr_grid))]
    corr_grid <- corr_grid[complete.cases(corr_grid), ]
  }

  annot_grid <- dcast(drug_cor_res, Gene ~ Drug, value.var = "Direction_Curated", fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Gene)
  annot_grid <- annot_grid[g2p_gene_order, intersect(g2p_drug_order, colnames(signif_grid))]
  annot_grid <- annot_grid[complete.cases(annot_grid), ]
  # annot_grid[annot_grid == "No"] <- ""
  # annot_grid[annot_grid == "Yes"] <- "G"

@@

```{r drug_spear_plots, eval=FALSE}
# Copy number
ccle_cn_cor <- unfactorize(filter(cn_cor, Dataset == "CCLE" & Gene %in% unique(g2p_amp$Gene)))
ccle_cn_ht <- MakeCorrHeatmap(drug_cor_res = ccle_cn_cor, row_names = TRUE, col_title = "CCLE", color_range = c(-0.2, 0, 0.2), radius = 0.1, legend_title = "CCLE Spearman correlation", metric = "CN")
png(filename = "./plots/manuscript/ccle_cn_heatmap.png", width = 3, height = 3, units = "in", res = 500)
draw(ccle_cn_ht, heatmap_legend_side = "top")
dev.off()

>>>>

Spearman correlations for gene expression and copy number:
```{r spear_calc, warning=FALSE, eval=FALSE}
# CRISPR
crispr_spear_cn <- transform(crispr %>% group_by(Gene) %>% drop_na(CN) %>% summarize(p = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$p.value, p_log10 = -log10(p), corr = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$estimate, Metric = "Copy Number"), Gene = reorder(Gene, corr))
c("white", "deepskyblue4")
crispr_spear_cn <- add_column(crispr_spear_cn, Drug = "CRISPR", .after = "Gene")
saveRDS(crispr_spear_cn, "./data_munging/rds/crispr_cn_cor.rds")

crispr_spear_ge <- transform(crispr %>% group_by(Gene) %>% drop_na(TPM) %>% summarize(p = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$p.value, p_log10 = -log10(p), corr = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$estimate, Metric = "Gene Expression"), Gene = reorder(Gene, corr))
crispr_spear_ge <- add_column(crispr_spear_ge, Drug = "CRISPR", .after = "Gene")
saveRDS(crispr_spear_ge, "./data_munging/rds/crispr_ge_cor.rds")

# CCLE
ccle_filt <- filter(ccle, Drug %in% g2p_drugs & Gene %in% g2p_genes)
ccle_filt$Gene <- factor(ccle_filt$Gene, ordered = FALSE)
ccle_filt <- unfactorize(ccle_filt)
ccle_ge_res <- ccle_filt %>% group_by(Drug, Gene, Drug_Gene) %>% drop_na(TPM) %>% summarize(GE_r = cor.test(y = Score, x = TPM, method = "spearman", use = "na.or.complete")$estimate, GE_p = cor.test(y = Score, x = TPM, method = "spearman", use = "na.or.complete")$p.value)
ccle_ge_res$InG2P <- ifelse(ccle_ge_res$Drug_Gene %in% unique(as.character(g2p_expr$Drug_Gene[!is.na(g2p_expr$Drug_Gene)])), "Yes", "No")
ccle_ge_res$Dataset <- "CCLE"

ccle_cn_res <- ccle_filt %>% group_by(Drug, Gene, Drug_Gene) %>% drop_na(CN) %>% summarize(CN_r = cor.test(y = Score, x = CN, method = "spearman", use = "na.or.complete")$estimate, CN_p = cor.test(y = Score, x = CN, method = "spearman", use = "na.or.complete")$p.value)
ccle_cn_res$InG2P <- ifelse(ccle_cn_res$Drug_Gene %in% unique(as.character(g2p_amp$Drug_Gene[!is.na(g2p_amp$Drug_Gene)])), "Yes", "No")
ccle_cn_res$Dataset <- "CCLE"

# CTRP
ctrp_filt <- filter(ctrp, Drug %in% g2p_drugs & Gene %in% g2p_genes)
ctrp_filt$Gene <- factor(ctrp_filt$Gene, ordered = FALSE)
ctrp_filt <- unfactorize(ctrp_filt)
ctrp_ge_res <- ctrp_filt %>% group_by(Drug, Gene, Drug_Gene) %>% drop_na(TPM) %>% summarize(GE_r = cor.test(y = Score, x = TPM, method = "spearman", use = "na.or.complete")$estimate, GE_p = cor.test(y = Score, x = TPM, method = "spearman", use = "na.or.complete")$p.value)
ctrp_ge_res$InG2P <- ifelse(ctrp_ge_res$Drug_Gene %in% unique(as.character(g2p_expr$Drug_Gene[!is.na(g2p_expr$Drug_Gene)])), "Yes", "No")
ctrp_ge_res$Dataset <- "CTRP"

ctrp_cn_res <- ctrp_filt %>% group_by(Drug, Gene, Drug_Gene) %>% drop_na(CN) %>% summarize(CN_r = cor.test(y = Score, x = CN, method = "spearman", use = "na.or.complete")$estimate, CN_p = cor.test(y = Score, x = CN, method = "spearman", use = "na.or.complete")$p.value)
ctrp_cn_res$InG2P <- ifelse(ctrp_cn_res$Drug_Gene %in% unique(as.character(g2p_amp$Drug_Gene[!is.na(g2p_amp$Drug_Gene)])), "Yes", "No")
ctrp_cn_res$Dataset <- "CTRP"

# GDSC
gdsc_filt <- filter(gdsc, Drug %in% g2p_drugs & Gene %in% g2p_genes)
gdsc_filt$Gene <- factor(gdsc_filt$Gene, ordered = FALSE)
gdsc_filt <- unfactorize(gdsc_filt)
gdsc_ge_res <- gdsc_filt %>% group_by(Drug, Gene, Drug_Gene) %>% drop_na(TPM) %>% summarize(GE_r = cor.test(y = Score, x = TPM, method = "spearman", use = "na.or.complete")$estimate, GE_p = cor.test(y = Score, x = TPM, method = "spearman", use = "na.or.complete")$p.value)
gdsc_ge_res$InG2P <- ifelse(gdsc_ge_res$Drug_Gene %in% unique(as.character(g2p_expr$Drug_Gene[!is.na(g2p_expr$Drug_Gene)])), "Yes", "No")
gdsc_ge_res$Dataset <- "GDSC"

gdsc_cn_res <- gdsc_filt %>% group_by(Drug, Gene, Drug_Gene) %>% drop_na(CN) %>% summarize(CN_r = cor.test(y = Score, x = CN, method = "spearman", use = "na.or.complete")$estimate, CN_p = cor.test(y = Score, x = CN, method = "spearman", use = "na.or.complete")$p.value)
gdsc_cn_res$InG2P <- ifelse(gdsc_cn_res$Drug_Gene %in% unique(as.character(g2p_amp$Drug_Gene[!is.na(g2p_amp$Drug_Gene)])), "Yes", "No")
gdsc_cn_res$Dataset <- "GDSC"

@@

```{r load_spear}
g2p_amp_crispr_4merge <- select(g2p_amp, Gene, Direction_Curated)

crispr_spear_cn <- readRDS("./data_munging/rds/crispr_cn_cor.rds")
crispr_spear_ge <- readRDS("./data_munging/rds/crispr_ge_cor.rds")

g2p_amp_4merge <- select(g2p_amp, Drug_Gene, Direction_Curated)
cn_cor <- readRDS("./data_munging/rds/drug_cn_cor.rds")
cn_cor <- merge(cn_cor, g2p_amp_4merge, by = "Drug_Gene", all.x = TRUE)

@@

```{r crispr_spear_eda, warning=FALSE}
crispr_spear_cn_plot <- ggplot(data = filter(crispr_spear_cn, Gene %in% g2p_dg_genes)) +
  geom_point(aes(x = Gene, y = corr, fill = p_log10), pch = 21, size = 5) +
  theme(legend.position = "bottom") +
  scale_fill_gradientn(colours = c("white", "green4"),
                       breaks = c(0, 4, 8, 12, 16, 20),
                       labels = c(0, 4, 8, 12, 16, 20),
                       limits = c(0, 21)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10)) +
  labs(title = "Copy Number",
       fill = expression(-log[10](p)),
       x = "Gene",
       y = expression("Spearman Correlation Coefficient"~(r[s])))

@@

```{r}
    signif_df <- filter(crispr_signif_g2p_g, Gene %in% unique(g2p_mut$Gene))
    signif_grid <- dcast(signif_df, Gene ~ Drug, value.var = "p_log10", fun.aggregate = mean, fill = 0)
    signif_grid <- signif_grid[match(g2p_gene_order, signif_grid$Gene),] %>% drop_na(Gene) %>% `rownames<-`(.[,1]) %>% select(-Gene)

```{r corr_heatmap_function}
MakeCorrHeatmap <- function(cor_res, row_names, col_title, color_range, legend_title, metric, radius, crispr) {
  if(toupper(metric) == "CN" | toupper(metric) == "COPY NUMBER") {
    if(crispr == TRUE) {

    }
    if(crispr == FALSE) {
      signif_grid <- dcast(cor_res, Gene ~ Drug, value.var = "CN_p", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Gene)
      signif_grid <- signif_grid[g2p_gene_order, intersect(g2p_drug_order, colnames(signif_grid))]
      signif_grid <- signif_grid[complete.cases(signif_grid), ]
      signif_grid[signif_grid >= 0.05] <- ""
      signif_grid[signif_grid != ""] <- "*"

      corr_grid <- dcast(cor_res, Gene ~ Drug, value.var = "CN_r", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Gene)
      corr_grid <- corr_grid[g2p_gene_order, intersect(g2p_drug_order, colnames(corr_grid))]
      corr_grid <- corr_grid[complete.cases(corr_grid), ]
    }
  }

  if(toupper(metric) == "GE" | toupper(metric) == "GENE EXPRESSION") {
    if(crispr == TRUE) {

    }
    if(crispr == FALSE) {
      signif_grid <- dcast(cor_res, Gene ~ Drug, value.var = "GE_p", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Gene)
      signif_grid <- signif_grid[g2p_gene_order, intersect(g2p_drug_order, colnames(signif_grid))]
      signif_grid <- signif_grid[complete.cases(signif_grid), ]
      signif_grid[signif_grid >= 0.05] <- ""
      signif_grid[signif_grid != ""] <- "*"

      corr_grid <- dcast(cor_res, Gene ~ Drug, value.var = "GE_r", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Gene)
      corr_grid <- corr_grid[g2p_gene_order, intersect(g2p_drug_order, colnames(corr_grid))]
      corr_grid <- corr_grid[complete.cases(corr_grid), ]
    }
  }
  if(crispr == TRUE) {
    annot_grid <- dcast(cor_res, Gene ~ Drug, value.var = "Direction_Curated", fill = 0)
    annot_grid <- annot_grid[match(g2p_gene_order, annot_grid$Gene),] %>% drop_na(Gene) %>% `rownames<-`(.[,1]) %>% select(-Gene)
    # annot_grid[annot_grid == "sensitivity"] <- "O"
    }
  if(crispr == FALSE) {
    annot_grid <- dcast(cor_res, Gene ~ Drug, value.var = "Direction_Curated", fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Gene)
    annot_grid <- annot_grid[g2p_gene_order, intersect(g2p_drug_order, colnames(signif_grid))]
    annot_grid <- annot_grid[complete.cases(annot_grid), ]
  }

@@

```{r drug_spear_plots, eval=FALSE}
# CRISPR
crispr_spear_cn <- unfactorize(filter(crispr_spear_cn, Gene %in% unique(g2p_amp$Gene)))
crispr_cn_ht <- MakeCorrHeatmap(drug_cor_res = crispr_spear_cn, row_names = TRUE, col_title = "CCLE", color_range = c(-0.2, 0, 0.2), radius = 0.1, legend_title = "CCLE Spearman correlation", metric = "CN")
png(filename = "./plots/manuscript/crispr_cn_heatmap.png", width = 3, height = 3, units = "in", res = 500)
draw(crispr_cn_ht, heatmap_legend_side = "top")
dev.off()

# Copy number
ccle_cn_cor <- unfactorize(filter(cn_cor, Dataset == "CCLE" & Gene %in% unique(g2p_amp$Gene)))
ccle_cn_ht <- MakeCorrHeatmap(drug_cor_res = ccle_cn_cor, row_names = TRUE, col_title = "CCLE", color_range = c(-0.2, 0, 0.2), radius = 0.1, legend_title = "CCLE Spearman correlation", metric = "CN")
png(filename = "./plots/manuscript/ccle_cn_heatmap.png", width = 3, height = 3, units = "in", res = 500)
draw(ccle_cn_ht, heatmap_legend_side = "top")
dev.off()
